---
title: "writing_functions"
output: github_document
date: "2023-10-26"
---

```{r}
library(tidyverse)
library(rvest)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%")

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis")

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Set seed for reproducibility. 

```{r}
set.seed(12345)
```

### Z score function

Z scores subtract the mean and divide by standard deviation.

```{r}
x_vec = rnorm(20, mean = 5, sd = .3)

x_vec
```

Compute Z scores for `x_vec`:

```{r}
(x_vec - mean(x_vec)) / sd(x_vec)
```

Write a function that takes the input (collection of numbers), output (collection of Z-scores), and the collection is the calculation of z-scores from a random sample seed. 

```{r}
z_score = function(x){
  
  z = (x - mean(x)) / sd(x)
  
  z
  
}
```

Check that this works. 

```{r}
z_score(x = x_vec)

z_score(x = rnorm(10, mean = 5))
```

^ the input needs to be the exact same everytime - be sure that you're not giving the input different names everytime. 

Keep checking. 

^ we wrote our function for a collection of numbers. since these are not a collection of numbers, these won't work in the function. 

Let's update the function to add some checks.

^ fix these errors before you can knit + commit + push to Github



```{r}
mean_and_sd = function(x) {
  
  if(!is.numeric(x)){
    stop("Argument should be numbers")
  } else if (length(x) < 2){
    stop("You need at least 2 numbers to get z scores")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

  tibble(
    mean = mean_x,
    sd = sd_x
  )
}
```

```{r}
mean_and_sd(x_vec)
```


### Start getting sample means and sample sds 

```{r}
x_vec = rnorm(n = 30, mean = 5, sd = .5)

tibble(
    mean = mean(x_vec),
    sd = sd(x_vec)
  )
```

^^?? why does the tibble not actually show up when I run the code ??????

Let's write a function that uses `n`, a true mean, and a true SD as inputs. 

```{r}
sim_mean_sd = function(n_obs, mu = 5, sigma = 1){
  
  x_vec = rnorm(n = n, mean = mu, sd = sigma)

  tibble(
    mean = mean(x_vec),
    sd = sd(x_vec)
  )
  
}

sim_mean_sd(n_obs = 3000, mu = 50)

sim_mean_sd(12, 24, 4)
```

^ review rnorm and population sample mean and pop sizes

^ ?? why doesn't the output show?


#### learning assessment - LoTR words

```{r}
lotr_load_and_tidy = function(path = "data/LotR_Words.xlsx", cell_range, movie_name){
  
  movie_df = 
    readxl::read_excel(path, range = cell_range) |>
    mutate(movie = movie_name) |> 
    janitor::clean_names() |> 
    pivot_longer(
      female:male,
      names_to = "sex",
      values_to = "words"
    ) |> 
    selex(movie, everything())
  
  movie_df
  
}

lotr_df = 
  bind_rows(
    lotr_load_and_tidy(cell_range = "B3:D6", movie_name = "fellowship_ring"),
    lotr_load_and_tidy(cell_range = "F3:H6", movie_name = "two_towers"),
    lotr_load_and_tidy(cell_range = "J3:L6", movie_name = "return_king")
  )
```

another example...NSDUH


```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_table = function(html, table_number, outcome_name) {
  
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent),
      name = table_name) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
}

nsduh_table(html = nsduh_html, table_number = 1, outcome_name = "marj")
nsduh_table(html = nsduh_html, table_number = 4, outcome_name = "cocaine")
```










